---
title: "1st assigment"
author: "Pep Porrà"
date: "May, 15th 2016"
output: html_document
---

## Basic requirements
```{r libraries, echo=TRUE, cache=TRUE}
library(knitr)
```

```{r setoptions, echo=TRUE, cache=TRUE}
opts_chunk$set(echo = FALSE, cache = TRUE)
```

## 1. Read data file

### Download file
```{r, echo=TRUE, cache=TRUE}
temp <- tempfile()
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", temp)
```

### unzip file
```{r, echo=TRUE, cache=TRUE}
unzip(temp)
unlink(temp)
```

### read file
```{r, echo=TRUE, cache=TRUE}
activity <- read.csv("activity.csv", header = TRUE, stringsAsFactors = FALSE)
```

### check data
```{r, echo=TRUE, cache=TRUE}
str(activity)
```

```{r, echo=TRUE, cache=TRUE}
summary(activity)
```

```{r, echo=TRUE, cache=TRUE}
length(unique(as.Date(activity$date)))
```

```{r, echo=TRUE, cache=TRUE}
data.frame( N = NROW(activity),
  N_data = NROW(activity[complete.cases(activity), ]),
  N_NA = NROW(activity) - NROW(activity[complete.cases(activity), ])
  )
```

#### days with NA activity

NA activity only applies to certain days and during the whole day
```{r, echo=TRUE, cache=TRUE}
df_na <- activity[!complete.cases(activity), ]
table(df_na$date)
```

Vector with days without any data
```{r, echo=TRUE, cache=TRUE}
na_dates <- names(table(df_na$date))
na_dates
```

Test
```{r, echo=TRUE, cache=TRUE}
2304 == 288 * length(na_dates)
```

## 2. Histogram of the total steps per day

### Data: total steps per day
```{r}
steps_day <- tapply(activity$steps, INDEX = factor(activity$date), sum)
```

### Histogram of steps per day

```{r, echo=TRUE, cache=TRUE}
hist(steps_day,
  xlab = "Number of steps per day",
  main = "Histogram of steps per day", breaks = 10)
```

## 3. mean and median of steps per day

```{r}
data.frame( mean = mean(steps_day, na.rm = TRUE),
  median = median(steps_day, na.rm = TRUE))
```

Be careful: Summary onñy takes 3 digits
```{r}
summary(steps_day, digits = 8)
```

## 4. Time series plot of the average number of steps taken

### Data: Average activity per interval (all days averaged)
```{r}
activity_clean <- activity[complete.cases(activity), ]
interval_steps <- tapply(activity_clean$steps, 
  INDEX = factor(activity_clean$interval), mean)
```

Notation for the interval is: h*100+min where h (0, 1, ..., 23) and min (0, 5, 10, ..., 55)

```{r}
str(factor(activity_clean$interval))
length(unique(activity_clean$interval))
head(table(activity_clean$interval), 15)
```

```{r}
plot(unique(activity_clean$interval), interval_steps, typ = "l",
  main = "Average steps per 5 min interval", 
  xlab = "5 minutes ordinal",
  ylab = "Average steps")
```

## 5. 5-minute interval with maximum number of steps

```{r}
cond <- interval_steps == max(interval_steps)
resp <- interval_steps[cond]
resp
resp_interval <- as.integer(names(resp))
resp_h <- resp_interval %/% 100
resp_min <- resp_interval %% 100
```

```{r}
cat(  paste0(
  "Maximum number of average steps (", 
  resp, 
  ") corresponds to 5-minutes interval ",
  resp_interval, " (", resp_h, ":", resp_min, ")")
)
```

## 6. Imputing missing data

We already obtained dates with missing data
```{r}
na_dates
```

We look first if there is a trend in the activity per day

```{r}
summary(lm(steps_day ~ as.Date(names(steps_day))))
```

It does not look that there is such a trend

```{r}
plot(as.Date(unique(activity$date)), steps_day,
  xlab = "date", ylab = "steps per day")
abline(v=as.Date(na_dates), col = 'red', lty = 2)
abline(reg = lm(steps_day ~ as.Date(names(steps_day)))
  , col = "green 4", lty = 4, lwd = 2)
```

We replace the missing values by the mean of the steps per 5-minutes interval averaged over all days.

```{r}
activity_complete <- activity
activity_complete[ activity_complete$date %in% na_dates, "steps"] <- 
  interval_steps
```

## 7. Histogram once all missing imputed

```{r}
hist(tapply(activity_complete$steps, INDEX = factor(activity_complete$date), sum),
  xlab = "Number of steps per day",
  main = "Histogram of steps per day (Missing inputed)", breaks = 10)
```


